The Heavy Photon Search Test Detector

### 2014-04-08 hps-testrun-nim.tex
Alphabetical author list. Consistent notion of HPS Test Run.

\input{author_list_hps_PAC39_compact}

@@

\begin{abstract}
The  Heavy Photon Search (HPS), an experiment to search for a hidden sector photon in fixed target 
electroproduction, is preparing for installation at the Thomas Jefferson National Accelerator Facility 
(JLab) in the Fall of 2014.  As the first stage of this project, the HPS Test Run apparatus was 
constructed and operated in 2012 to demonstrate the experiment's technical feasibility and to  confirm 
that the trigger rates and occupancies are as expected. This paper describes the Test Run apparatus 
and readout electronics and its performance. In this setting, a heavy photon can be identified as a 
narrow\ee  mass peak above the trident background or as a decay vertex displaced from
the production target, so charged particle tracking and vertexing are needed for its detection. In the 
HPS Test Run, charged particles are measured
with a compact forward silicon microstrip vertex/tracker inside a dipole magnet. Electromagnetic 
showers are detected in a PbW0$_{4}$ 
crystal calorimeter situated behind the magnet used to trigger the experiment and identify 
electrons and positrons. Both detectors are placed close to the 
beam line and split top-bottom. This arrangement  provides sensitivity to low-mass heavy photons,  
allows clear  passage of the unscattered beam, 
and avoids the spray of degraded electrons coming from the target. 
The discrimination between prompt and displaced \ee pairs requires the first layer of silicon 
sensors be placed only 10~cm downstream of the target. The expected signal is small, and the 
trident background huge, so the experiment requires very large statistics. Accordingly, the
HPS Test Run utilizes  high-rate readout and data acquisition electronics and a fast trigger to
perfectly exploit the essentially 100\% duty cycle of the CEBAF accelerator.
\end{abstract}

@@

%% main text
\section{Introduction}
\label{introduction}
The heavy photon (\Aprime{}), aka a "hidden sector" or "dark" photon, is a particle with mass 
$10-1000$~MeV which couples weakly to electric charge by mixing with the Standard Model 
photon~\cite{Holdom:1985ag}. Consequently, it can be radiated by electrons and subsequently decay 
into \ee pairs, albeit at rates far below those of QED trident processes. Heavy photons have been 
suggested by numerous Beyond Standard Model theories `\cite{Essig:2013lka}, to explain the 
discrepancy 
between theory and experiment of the muon's $g-2$ ~\cite{Pospelov:2008zw}, and as a possible 
explanation of recent astrophysical anomalies.  Heavy photons  couple directly to  hidden sector 
particles with "dark" or "hidden sector" charge; these particles
could constitute all or some of the dark matter~\cite{ArkaniHamed:2008qn}. 
Current phenomenology highlights the $20-1000$~MeV/c$^{2}$ mass range, and 
suggests that  the coupling to electric charge, $\epsilon$e, has $\epsilon$ in the range of 
$10^{-3} -10^{-5}$. [REFERENCE]. This range of parameters makes \Aprime{} searches viable in 
medium energy fixed target electroproduction~\cite{Bjorken:2009mm}, but requires large data sets and 
good mass resolution to identify a small mass peak above the copious QED background. At small 
couplings, \Aprime{} become long-lived, so detection of a displaced decay vertex can reject the prompt 
QED background and boost experimental sensitivity.

The HPS experiment~\cite{HPS_proposal_2010} uses both invariant mass and secondary
vertex signatures to search for A's. It uses  a $\approx 1$~m long silicon tracking and vertexing detector
inside a dipole magnet to measure charged particle trajectories and a fast electromagnetic calorimeter 
just downstream of the magnet to provide a trigger and identify electrons. The experiment utilizes very 
high rate front end electronics and runs at high trigger rates (up to 50kHz), exploiting the 100\% duty 
cycle of the Thomas Jefferson National Accelerator Facility (JLab) CEBAF accelerator to accumulate 
the needed statistics.

The HPS Test apparatus, a simplified version of the HPS experiment, was proposed and 
approved at Thomas Jefferson National Accelerator Facility (JLab) as the first stage of HPS. 
Its purposes included demonstrating that the apparatus and data acquisition systems are 
technically feasible and  the trigger rates and occupancies to be  encountered in electron-beam 
running are as simulated. Given dedicated running time with electron beams, the HPS Test Run 
apparatus is capable of searching for heavy photons in unexplored regions of parameter space.    
The key design criteria for HPS and the HPS Test Run are the same:
\begin{itemize}
\item large and uniform acceptance in the forward region close to the beam in order to catch boosted 
\Aprime{} decay products,
\item beam passage through the apparatus in vacuum, to eliminate direct interactions with the 
detector and minimize beam gas interactions, 
\item a flexible, redundant and efficient trigger selecting electron and positron pairs at rates up to 
50~kHz,
\item excellent track reconstruction efficiency for electrons and positrons,
\item good angular and momentum resolution to reconstruct invariant mass precisely,
\item excellent vertex resolution to discriminate displaced \Aprime{} decays from prompt QED 
backgrounds,
\item high rate electronics with excellent timing resolution to minimize out of time backgrounds,
\item data handling rates of 100~MB/s to permanent storage,
\item detector components that can survive and efficiently operate in a high radiation environment 
with local doses exceeding $100$~Mrad.
\end{itemize}

The Test apparatus was installed on April 19, 2012, and ran parasitically in the photon beam of the HDice 
experiment~\cite{HDice} until May 18. The JLab run schedule 
precluded any dedicated electron beam running, but the HPS Test Run was allowed a short and 
valuable dedicated run at the end of scheduled CEBAF running. This final running provided
enough data to demonstrate the functionality of the apparatus, document its performance, and
explore trigger rates, as shown below. 
This paper reviews the HPS Test apparatus; documents the performance of the trigger, 
data acquisition, silicon tracking detector, and the electromagnetic calorimeter at the level assumed 
in calculating the physics reach of the HPS experiment.

This paper reviews the HPS Test apparatus; documents the performance of the trigger, 
data acquisition, silicon tracking detector, and the electromagnetic calorimeter at the level assumed 
in calculating the physics reach of the HPS experiment.

\section{Detector Overview}
\label{detector}
{\small
\begin{figure*}[t]
\begin{center}
    \includegraphics[width=0.8\textwidth]{figures/hps_testrun_rendering}
\caption{Rendering of the HPS Test apparatus installed on the beam line.}
\label{fig:testrundetector}
\end{center}
\end{figure*}
}
The HPS Test apparatus was designed to run in Hall~B at JLab using the CEBAF electron beam, a 
499~MHz beam, at energies of 2.2 and 6.6~GeV and currents between 200 and 600~nA.  
The overall design of the experiment follows from the kinematics of \Aprime{} production which 
typically results in a final state particle within a few degrees of the incoming beam, especially at low 
$m_{\Aprime}$. Detectors must therefore be placed close to the beam. 
The intense electron beam enlarges downstream after multiple scattering in the target and electrons 
which have radiated in the target disperse horizontally in the field of the analyzing magnet. Together 
they constitute a "wall of flame" which must be completely avoided. Accordingly, 
the apparatus is split vertically to avoid a "dead zone", the region within $\pm 15$~mrad of the beam 
plane. In addition, the beam is transported in vacuum through the tracker to minimize beam-gas 
interaction backgrounds. Even with these precautions, the occupancies of sensors near the beam 
plane are high, dominated by the
multiple Coulomb scattering of the primary beam, so high rate detectors, a fast trigger, 
and excellent time tagging are required to minimize their impact.  
The trigger comes from a highly-segmented  PbWO$_{4}$ crystal calorimeter located just downstream 
of the dipole magnet. 
 %\begin{multicols}{2}
\begin{center}
\begin{table*}[t]
{\small
\caption{Overview of the coverage, segmentation and performance of the HPS Test detector. }
\begin{tabular}{lccccccc}
\hline 
System & Coverage & \# channels & ADC & Time resolution & \# layers & Segmentation & Performance \\
 & (mrad) &  & (bit) & (ns) &  &  &  \\
\hline
SVT & $15<\theta_{y} < XXXX$ & 12800 & 14 & $\approx 2$~ns & 5  & $\approx 120~\mu$m $r-\phi$ & $\sigma_{d0,y}  \approx 100~\mu$m \\
& &  &  &  & (stereo layers) & $\approx 6~\mu$m $z$ & $\sigma_{d0,x} \approx 300~\mu$m \\
& &  &  &  &  &  & $\sigma_{d0,z}\approx 1$~mm \\
\hline
ECal & $15<\theta_{y} < XXXX$ & 442 & 12 & 4~ns & 1 & $1.3\times1.3$~cm$^2$  & $\sigma(E)/E \approx 4.5\%$ \\ 
 &  &  &  &  &  & $1.6\times1.6$~cm$^2$  &  \\ 
\hline
\end{tabular}
\label{tab:detector-overview}
}
\end{table*}
\end{center}
%\end{multicols}
A rendering of the apparatus installed on the beam line is shown in 
Fig.~\ref{fig:testrundetector} and an overview of the coverage, segmentation and performance is 
given in Tab.~\ref{tab:detector-overview}.

The silicon tracking and vertexing detector for HPS Test, or SVT, resides in a vacuum 
chamber inside the Pair Spectrometer (PS) dipole magnet in Hall B at JLab. The SVT has five 
measurement stations, or ``layers,'' beginning10 cm downstream of the target. Each layer 
comprises a pair of closely-spaced silicon microstrip sensors responsible for measuring a single 
coordinate, or ``view.'' Introduction of a small (50 or 100~mrad) stereo angle between the two 
sensors of each layer provides three-dimensional tracking and vertexing throughout the acceptance 
of the detector. In order to accommodate the dead zone, the 
SVT is built in two halves that are approximately mirror reflections of one another about the plane of the 
nominal electron beam.  Each layer in one half is supported on a common support plate with 
independent cooling and readout.

@@

\section{The HPS Test Run Beamline}
The HPS Test Run studied multiple Coulomb scattering of electrons and positrons from 
bremsstrahlung photons produced in the Hall B tagged photon facility. 
Figure~\ref{fig:hpstest_layout} shows the layout of 
the setup on the beam line. The SVT was installed inside the Hall B pair 
spectrometer magnet vacuum chamber with the ECal mounted downstream of it. Both the 
SVT and the ECal were retracted off the beam plane compared to nominal electron beam runing to 
allow clean passage of the photon beam through the system. 
\begin{figure}[]
    \includegraphics[width=0.5\textwidth]{figures/HPS_dimensions}
\caption{\small{Layout of the HPS parasitic run.} }
\label{fig:hpstest_layout}
\end{figure}

@@

 Data was taken on three different converter thicknesses with photon fluxes between 
 1.1-2.6$\times10^{8}$/s ($0.55<$E$_\gamma<5.5$~GeV) at beam currents varying between 
 30-90~nA and repeated with the reverse field setting of the PS dipole magnet.
\begin{table}[]
\begin{center}
{\small
\begin{tabular}{|c|c|c|}
\hline
Converter thickn. & Duration &  $e^-$ on radiator \\
 (\%$X_0$) & (s) & ($\mu$C)    \\   
\hline
%\hline
1.6   & 911 &   24.4 \\ %24385.9     \\ %27 nA
%\hline
0.18   & 2640 &   193.5 \\ % 193508.9  \\  %73 nA
%\hline
0.45  & 2149 &     140.7 \\ %  140709.9  \\ %65.5 nA
%\hline
0    & 1279  &   88.1 \\ %88079.6  \\
\hline
\end{tabular}
}
\caption{Measured integrated currents for the dedicated photon runs.}
\label{tab:currents}
\end{center}
\end{table}
The photon beam line during the Test Run produced a relatively large fraction of pairs 
originating upstream of the converter position. This contribution was measured during data taking 
with ``empty'' converter runs i.e. removing the converter but with all other conditions the same. 
The runs taken during the time dedicated to HPS Test is summarized in Tab.~\ref{tab:currents}.

@@

The design of the SVT is primarily driven by direct physics requirements and constraints from the 
environment at the interaction region. The \Aprime{} decay products 
have momenta in the range of 1~GeV, so multiple scattering dominates mass and vertexing 
uncertainties for any possible material budget, so the SVT must minimize the amount of 
material in the tracking volume. The signal yield for long-lived \Aprime{} is very small, so 
the rejection of prompt vertexes must be exceedingly pure, on the order of $10^{-7}$, in order to 
eliminate all prompt backgrounds. To achieve the required vertexing performance the first layer of the 
SVT must be placed no more than about 10 cm downstream of the target. At that distance, it is found 
that the active region of a sensor can be placed as close to 1.5 mm from the center of the beam, 
defining the 15 mrad ``dead zone" mentioned previously, to maximize low-mass \Aprime{} acceptance 
with decay products nearly collinear with the beam axis. At the edge of this ``dead zone," the 
radiation dose approaches $10^{15}$ electrons/cm$^2$/month, or roughly 
$3 \times 10^{14}$~\fluenceunit{}/month~\cite{Rashevskaya:2002nd}, 
requiring the sensors to be actively cooled.
% and~\ref{fig:rates}.
%===================
%\begin{figure}[]
%\includegraphics[width=7cm]{figures/radiation}
%\label{fig:radiation}
%\caption{{\small Simulation of the electron fluence 10~cm downstream of the target in the active region 
%of the SVT ($|Y|>0.15~$mm) looking downstream; electrons bend towards $+x$.}}
%\end{figure}
Meanwhile, very low-energy delta rays from 
beam-gas interactions multiply the density of background hits, so the SVT must operate inside the 
beam vacuum.  Finally, in order to protect the sensors, the detector must be movable so that it can be 
retracted during periods of uncertain beam conditions.

\subsection{Layout}
The layout of the SVT is summarized in Tab.~\ref{tab:trk} and rendered in Fig.~\ref{fig:tracker_model}. 
Each of the layers is comprised of a pair of closely-spaced silicon microstrip sensors mounted back-
to-back to form a module. A 100~mrad stereo angle is used in the first three layers to provide higher-
resolution 3D space points for vertexing.  Using 50~mrad in the last two layers breaks the tracking 
degeneracy of having five identical layers and minimizes fakes from ghost hits to improve pattern 
recognition. Altogether, the SVT has 20 sensors for a total of 12780 readout channels. 
%=======================
\begin{center}
\begin{table}[ht]
{\footnotesize
\begin{tabular}{lccccc}   
\hline \hline 
    Layer & 1 & 2 & 3 & 4 & 5 \\      
\hline
    $z$ from target (cm)  & 10 & 20 & 30 & 50 & 70  \\ 
    Stereo angle (mrad)  & 100 & 100 & 100 & 50 & 50 \\ 
    Bend res. ($\mu$m)  & $\approx$60 & $\approx$60 & $\approx$60 & $\approx$120 & $\approx$120  \\ 
    Non-bend res. ($\mu$m)  & $\approx$6 & $\approx$6 & $\approx$6 & $\approx$6 & $\approx$6  \\ 
    \# of sensors  & 4 & 4 & 4 & 4 & 4  \\ 
    Dead zone (mm) & $\pm1.5$  & $\pm3.0$  & $\pm4.5$  & $\pm7.5$  & $\pm10.5$  \\ 
    Power cons. (W) & 6.9 & 6.9 & 6.9 & 6.9 & 6.9 \\
\hline \hline
\end{tabular}
\caption{\small Layout of the SVT.}
}
\label{tab:trk}
\vspace*{-15mm}
\end{table}
\end{center}
%=======================
\begin{center}
\begin{figure}[htp]
\includegraphics[width=7cm]{figures/HPS_nochamber}
\caption{\small A rendering of the SVT showing the modules on their support plates held by the 
hinged C-support on the left and the motion levers on the right. The sensors are shown in red and the 
hybrid readout boards in green. The beam enters from the right through a vacuum box with flanges 
for services. }
\label{fig:tracker_model}
%\vspace*{-5mm}
\end{figure}
\end{center}
%=======================
The SVT is built in two separate halves that are mirror reflections of one another about the plane of 
the nominal electron beam.  Each half consists of five modules mounted on a support plate that 
provides services to the modules and allows them to be moved as a group relative to the dead zone. 
The two halves of the tracker are connected to hinges mounted on a C-shaped support just beyond 
the last layer that defines the nominal spacing between the upper and lower halves of the tracker.  A 
shaft attached to each support plate in front of layer 1 extends upstream and connects to a linear shift 
that transfers motion into the vacuum box through bellows to open and close the two halves around 
the dead zone. The C-support is mounted to an aluminum baseplate that defines the position of the 
SVT with respect to the vacuum chamber. Figure~\ref{fig:tracker_halves} shows a photograph of both 
completed detector halves prior to final assembly. 
%=======================
\begin{figure}[htp]
\begin{center}
    \includegraphics[width=7cm]{figures/2012-101-PHOTON-DETECTOR-001}
\caption{\small{Both halves of the HPS Test SVT after final assembly at SLAC.  The cooling manifolds and 
integrated cable runs are clearly seen.} }
\label{fig:tracker_halves}
\end{center}
%\vspace*{-5mm}
\end{figure}
%=======================

\subsection{Components}
The sensors for the SVT are $p+$-on-$n$, single sided, AC coupled, polysilicon-biased microstrip 
sensors fabricated on $<100>$ silicon and have 30 (60) micron sense (readout) pitch over their 
$4\times10$~cm$^2$ surface. This sensor technology was selected to match the requirement of 
a $<1$\% $X_0$ per layer, single-hit resolution better than 50~$\mu$m and tolerant of a radiation 
dose of approximately $1.5\times10^{14}$~\fluenceunit{} for a six month run. The sensors 
were purchased from the Hamamatsu Photonics Corporation for the cancelled 
Run 2b upgrade of the D\O~experiment~\cite{Denisov:2001aa} which satisfied the requirement that 
the technology must be mature and available within the time and budget constraints.

@@

During the Test Run, both halves were held in place by four vertical bars attached to an above rail, 
placing the front face of the crystals 147~cm from the upstream edge of the magnet and with a 
8.7~cm gap between the innermost edge of the crystals in the two halves.
{\small
\begin{figure}[]
\begin{center}
\includegraphics[width=0.45\textwidth]{figures/ECal}
%\includegraphics[width=0.45\textwidth]{figures/ecal_picture_beamline}
\caption{Rendered layout view of the ECal looking downstream.}
\label{fig:ecal}
\end{center}
\end{figure}
}

>>>>

\input{author_list_hps_PAC39_compact_alphabetical_order}

@@

\begin{abstract}
The  Heavy Photon Search (HPS), an experiment to search for a hidden sector photon in fixed target 
electroproduction, is preparing for installation at the Thomas Jefferson National Accelerator Facility 
(JLab) in the Fall of 2014.  As the first stage of this project, the HPS Test Run apparatus was 
constructed and operated in 2012 to demonstrate the experiment's technical feasibility and to  confirm 
that the trigger rates and occupancies are as expected. This paper describes the HPS Test Run apparatus 
and readout electronics and its performance. In this setting, a heavy photon can be identified as a 
narrow \ee{}  mass peak above the trident background or as a decay vertex displaced from
the production target, so charged particle tracking and vertexing are needed for its detection. In the 
HPS Test Run, charged particles are measured
with a compact forward silicon microstrip vertex/tracker inside a dipole magnet. Electromagnetic 
showers are detected in a PbW0$_{4}$ 
crystal calorimeter situated behind the magnet used to trigger the experiment and identify 
electrons and positrons. Both detectors are placed close to the 
beam line and split top-bottom. This arrangement  provides sensitivity to low-mass heavy photons,  
allows clear  passage of the unscattered beam, 
and avoids the spray of degraded electrons coming from the target. 
The discrimination between prompt and displaced \ee{} pairs requires the first layer of silicon 
sensors be placed only 10~cm downstream of the target. The expected signal is small, and the 
trident background huge, so the experiment requires very large statistics. Accordingly, the
HPS Test Run utilizes  high-rate readout and data acquisition electronics and a fast trigger to
perfectly exploit the essentially 100\% duty cycle of the CEBAF accelerator.
\end{abstract}

@@

%% main text
\section{Introduction}
\label{introduction}
The heavy photon (\Aprime{}), aka a "hidden sector" or "dark" photon, is a particle with mass 
$10-1000$~MeV which couples weakly to electric charge by mixing with the Standard Model 
photon~\cite{Holdom:1985ag}. Consequently, it can be radiated by electrons and subsequently decay 
into \ee{} pairs, albeit at rates far below those of QED trident processes. Heavy photons have been 
suggested by numerous beyond Standard Model theories~\cite{Essig:2013lka}, to explain the 
discrepancy 
between theory and experiment of the muon's $g-2$ ~\cite{Pospelov:2008zw}, and as a possible 
explanation of recent astrophysical anomalies.  Heavy photons  couple directly to  hidden sector 
particles with "dark" or "hidden sector" charge; these particles
could constitute all or some of the dark matter~\cite{ArkaniHamed:2008qn}. 
Current phenomenology highlights the $20-1000$~MeV/c$^{2}$ mass range, and 
suggests that  the coupling to electric charge, $\epsilon$e, has $\epsilon$ in the range of 
$10^{-3} -10^{-5}$. This range of parameters makes \Aprime{} searches viable in 
medium energy fixed target electroproduction~\cite{Bjorken:2009mm}, but requires large data sets and 
good mass resolution to identify a small mass peak above the copious QED background. At small 
couplings, \Aprime{} become long-lived, so detection of a displaced decay vertex can reject the prompt 
QED background and boost experimental sensitivity.

The HPS experiment~\cite{HPS_proposal_2010} uses both invariant mass and secondary
vertex signatures to search for \Aprime{}. It uses  a $\approx 1$~m long silicon tracking and vertexing detector
inside a dipole magnet to measure charged particle trajectories and a fast electromagnetic calorimeter 
just downstream of the magnet to provide a trigger and identify electrons. The experiment utilizes very 
high rate front end electronics and runs at high trigger rates (up to 50kHz), exploiting the 100\% duty 
cycle of the Thomas Jefferson National Accelerator Facility (JLab) CEBAF accelerator to accumulate 
the needed statistics.

The HPS Test Run apparatus, a simplified version of the HPS experiment, was proposed and 
approved at JLab as the first stage of HPS. 
Its purposes included demonstrating that the apparatus and data acquisition systems are 
technically feasible and  the trigger rates and occupancies to be encountered in electron-beam 
running are as simulated. Given dedicated running time with electron beams, the HPS Test Run 
apparatus is capable of searching for heavy photons in unexplored regions of parameter space.    
The key design criteria for HPS and the HPS Test Run are the same:
\begin{itemize}
\item large and uniform acceptance in the forward region close to the beam in order to catch boosted 
\Aprime{} decay products,
\item beam passage through the apparatus in vacuum, to eliminate direct interactions with the 
detector and minimize beam gas interactions, 
\item a flexible, redundant and efficient trigger selecting electron and positron pairs at rates up to 
50~kHz,
\item excellent track reconstruction efficiency for electrons and positrons,
\item good angular and momentum resolution to reconstruct invariant mass precisely,
\item excellent vertex resolution to discriminate displaced \Aprime{} decays from prompt QED 
backgrounds,
\item high rate electronics with excellent timing resolution to minimize out of time backgrounds,
\item data handling rates of 100~MB/s to permanent storage,
\item detector components that can survive and efficiently operate in a high radiation environment 
with local doses exceeding 100~Mrad.
\end{itemize}

The HPS Test Run apparatus was installed on April 19, 2012, and ran parasitically in the photon beam of the HDice 
experiment~\cite{HDice} until May 18. The JLab run schedule 
precluded any dedicated electron beam running, but the HPS Test Run was allowed a short and 
valuable dedicated run at the end of scheduled CEBAF running. This final running provided
enough data to demonstrate the functionality of the apparatus, document its performance, and
explore trigger rates, as shown below. 
This paper reviews the HPS Test Run apparatus; documents the performance of the trigger, 
data acquisition, silicon tracking detector, and the electromagnetic calorimeter at the level assumed 
in calculating the physics reach of the HPS experiment.

This paper reviews the HPS Test Run apparatus; documents the performance of the trigger, 
data acquisition, silicon tracking detector, and the electromagnetic calorimeter at the level assumed 
in calculating the physics reach of the HPS experiment.

\section{Detector Overview}
\label{detector}
{\small
\begin{figure*}[t]
\begin{center}
    \includegraphics[width=0.8\textwidth]{figures/hps_testrun_rendering}
\caption{Rendering of the HPS Test Run apparatus installed on the beam line.}
\label{fig:testrundetector}
\end{center}
\end{figure*}
}
The HPS Test Run apparatus was designed to run in Hall~B at JLab using the CEBAF electron beam, a 
499~MHz beam, at energies of 2.2 and 6.6~GeV and currents between 200 and 600~nA.  
The overall design of the experiment follows from the kinematics of \Aprime{} production which 
typically results in a final state particle within a few degrees of the incoming beam, especially at low 
$m_{\Aprime{}}$. Detectors must therefore be placed close to the beam. 
The intense electron beam enlarges downstream after multiple scattering in the target and electrons 
which have radiated in the target disperse horizontally in the field of the analyzing magnet. Together 
they constitute a "wall of flame" which must be completely avoided. Accordingly, 
the apparatus is split vertically to avoid a "dead zone", the region within $\pm 15$~mrad of the beam 
plane. In addition, the beam is transported in vacuum through the tracker to minimize beam-gas 
interaction backgrounds. Even with these precautions, the occupancies of sensors near the beam 
plane are high, dominated by the
multiple Coulomb scattering of the primary beam, so high rate detectors, a fast trigger, 
and excellent time tagging are required to minimize their impact.  
The trigger comes from a highly-segmented  PbWO$_{4}$ crystal calorimeter located just downstream 
of the dipole magnet. 
 %\begin{multicols}{2}
\begin{center}
\begin{table*}[t]
{\small
\caption{Overview of the coverage, segmentation and performance of the HPS Test Run detector. }
\begin{tabular}{lccccccc}
\hline 
System & Coverage & \# channels & ADC & Time resolution & \# layers & Segmentation & Performance \\
 & (mrad) &  & (bit) & (ns) &  &  &  \\
\hline
SVT & $15<\theta_{y} < XXXX$ & 12800 & 14 & $\approx 2$~ns & 5  & $\approx 120~\mu$m $r-\phi$ & $\sigma_{d0,y}  \approx 100~\mu$m \\
& &  &  &  & (stereo layers) & $\approx 6~\mu$m $z$ & $\sigma_{d0,x} \approx 300~\mu$m \\
& &  &  &  &  &  & $\sigma_{d0,z}\approx 1$~mm \\
\hline
ECal & $15<\theta_{y} < XXXX$ & 442 & 12 & 4~ns & 1 & $1.3\times1.3$~cm$^2$  & $\sigma(E)/E \approx 4.5\%$ \\ 
 &  &  &  &  &  & $1.6\times1.6$~cm$^2$  &  \\ 
\hline
\end{tabular}
\label{tab:detector-overview}
}
\end{table*}
\end{center}
%\end{multicols}
A rendering of the apparatus installed on the beam line is shown in 
Fig.~\ref{fig:testrundetector} and an overview of the coverage, segmentation and performance is 
given in Tab.~\ref{tab:detector-overview}.

The silicon tracking and vertexing detector for HPS Test Run, or SVT, resides in a vacuum 
chamber inside the Pair Spectrometer (PS) dipole magnet in Hall B at JLab. The SVT has five 
measurement stations, or ``layers,'' beginning 10 cm downstream of the target. Each layer 
comprises a pair of closely-spaced silicon microstrip sensors responsible for measuring a single 
coordinate, or ``view''. Introduction of a small (50 or 100~mrad) stereo angle between the two 
sensors of each layer provides three-dimensional tracking and vertexing throughout the acceptance 
of the detector. In order to accommodate the dead zone, the 
SVT is built in two halves that are approximately mirror reflections of one another about the plane of the 
nominal electron beam.  Each layer in one half is supported on a common support plate with 
independent cooling and readout.

@@

\section{The HPS Test Run Beamline}
The HPS Test Run studied multiple Coulomb scattering of electrons and positrons from 
bremsstrahlung photons produced in the Hall B tagged photon facility. 
Figure~\ref{fig:hpstest_layout} shows the layout of 
the setup on the beam line. The SVT was installed inside the Hall B pair 
spectrometer magnet vacuum chamber with the ECal mounted downstream of it. Both the 
SVT and the ECal were retracted off the beam plane compared to nominal electron beam running to 
allow clean passage of the photon beam through the system. 
\begin{figure}[]
    \includegraphics[width=0.5\textwidth]{figures/HPS_dimensions}
\caption{\small{Layout of the HPS parasitic run.} }
\label{fig:hpstest_layout}
\end{figure}

@@

 Data was taken on three different converter thicknesses with photon fluxes between 
 0.4-1.3$\times10^{8}$/s ($0.55<$E$_\gamma<5.5$~GeV) at beam currents varying between 
 30-90~nA and repeated with the reverse field setting of the PS dipole magnet.
\begin{table}[]
\begin{center}
{\small
\begin{tabular}{|c|c|c|}
\hline
Converter thickn. & Duration &  $e^-$ on radiator \\
 (\%$X_0$) & (s) & ($\mu$C)    \\   
\hline
%\hline
1.6   & 911 &   24.4 \\ %24385.9     \\ %27 nA
%\hline
0.18   & 2640 &   193.5 \\ % 193508.9  \\  %73 nA
%\hline
0.45  & 2149 &     140.7 \\ %  140709.9  \\ %65.5 nA
%\hline
0    & 1279  &   88.1 \\ %88079.6  \\
\hline
\end{tabular}
}
\caption{Measured integrated currents for the dedicated photon runs.}
\label{tab:currents}
\end{center}
\end{table}
The photon beam line during the HPS Test Run produced a relatively large fraction of pairs 
originating upstream of the converter position. This contribution was measured during data taking 
with ``empty'' converter runs i.e. removing the converter but with all other conditions the same. 
The runs taken during the time dedicated to HPS Test Run is summarized in Tab.~\ref{tab:currents}.

@@

The design of the SVT is primarily driven by direct physics requirements and constraints from the 
environment at the interaction region. The \Aprime{} decay products 
have momenta in the range of 1~GeV, so multiple scattering dominates mass and vertexing 
uncertainties for any possible material budget, so the SVT must minimize the amount of 
material in the tracking volume. The signal yield for long-lived \Aprime{} is very small, so 
the rejection of prompt vertexes must be exceedingly pure, on the order of $10^{-7}$, in order to 
eliminate all prompt backgrounds. To achieve the required vertexing performance the first layer of the 
SVT must be placed no more than about 10~cm downstream of the target. At that distance, it is found 
that the active region of a sensor can be placed as close to 1.5~mm from the center of the beam, 
defining the 15~mrad ``dead zone" mentioned previously, to maximize low-mass \Aprime{} acceptance 
with decay products nearly collinear with the beam axis. At the edge of this ``dead zone," the 
radiation dose approaches $10^{15}$~electrons/cm$^2$/month, or roughly 
$3 \times 10^{14}$~\fluenceunit{}/month~\cite{Rashevskaya:2002nd}, 
requiring the sensors to be actively cooled.
% and~\ref{fig:rates}.
%===================
%\begin{figure}[]
%\includegraphics[width=7cm]{figures/radiation}
%\label{fig:radiation}
%\caption{{\small Simulation of the electron fluence 10~cm downstream of the target in the active region 
%of the SVT ($|Y|>0.15~$mm) looking downstream; electrons bend towards $+x$.}}
%\end{figure}
Meanwhile, very low-energy delta rays from 
beam-gas interactions multiply the density of background hits, so the SVT must operate inside the 
beam vacuum.  Finally, in order to protect the sensors, the detector must be movable so that it can be 
retracted during periods of uncertain beam conditions.

\subsection{Layout}
The layout of the SVT is summarized in Tab.~\ref{tab:trk} and rendered in Fig.~\ref{fig:tracker_model}. 
Each of the layers is comprised of a pair of closely-spaced silicon microstrip sensors mounted back-
to-back to form a module. A 100~mrad stereo angle is used in the first three layers to provide higher-
resolution 3D space points for vertexing.  Using 50~mrad in the last two layers breaks the tracking 
degeneracy of having five identical layers and minimizes fakes from ghost hits to improve pattern 
recognition. Altogether, the SVT has 20 sensors for a total of 12780 readout channels. 
%=======================
\begin{center}
\begin{table}[ht]
{\footnotesize
\begin{tabular}{lccccc}   
\hline \hline 
    Layer & 1 & 2 & 3 & 4 & 5 \\      
\hline
    $z$ from target (cm)  & 10 & 20 & 30 & 50 & 70  \\ 
    Stereo angle (mrad)  & 100 & 100 & 100 & 50 & 50 \\ 
    Bend res. ($\mu$m)  & $\approx$60 & $\approx$60 & $\approx$60 & $\approx$120 & $\approx$120  \\ 
    Non-bend res. ($\mu$m)  & $\approx$6 & $\approx$6 & $\approx$6 & $\approx$6 & $\approx$6  \\ 
    \# of sensors  & 4 & 4 & 4 & 4 & 4  \\ 
    Dead zone (mm) & $\pm1.5$  & $\pm3.0$  & $\pm4.5$  & $\pm7.5$  & $\pm10.5$  \\ 
    Power cons. (W) & 6.9 & 6.9 & 6.9 & 6.9 & 6.9 \\
\hline \hline
\end{tabular}
\caption{\small Layout of the SVT.}
}
\label{tab:trk}
\vspace*{-15mm}
\end{table}
\end{center}
%=======================
\begin{center}
\begin{figure}[htp]
\includegraphics[width=7cm]{figures/HPS_nochamber}
\caption{\small A rendering of the SVT showing the modules on their support plates held by the 
hinged C-support on the left and the motion levers on the right. The sensors are shown in red and the 
hybrid readout boards in green. The beam enters from the right through a vacuum box with flanges 
for services. }
\label{fig:tracker_model}
%\vspace*{-5mm}
\end{figure}
\end{center}
%=======================
The SVT is built in two separate halves that are mirror reflections of one another about the plane of 
the nominal electron beam.  Each half consists of five modules mounted on a support plate that 
provides services to the modules and allows them to be moved as a group relative to the dead zone. 
The two halves of the tracker are connected to hinges mounted on a C-shaped support just beyond 
the last layer that defines the nominal spacing between the upper and lower halves of the tracker.  A 
shaft attached to each support plate in front of layer 1 extends upstream and connects to a linear shift 
that transfers motion into the vacuum box through bellows to open and close the two halves around 
the dead zone. The C-support is mounted to an aluminum baseplate that defines the position of the 
SVT with respect to the vacuum chamber. Figure~\ref{fig:tracker_halves} shows a photograph of both 
completed detector halves prior to final assembly. 
%=======================
\begin{figure}[htp]
\begin{center}
    \includegraphics[width=7cm]{figures/2012-101-PHOTON-DETECTOR-001}
\caption{\small{Both halves of the HPS Test Run SVT after final assembly at SLAC.  The cooling manifolds and 
integrated cable runs are clearly seen.} }
\label{fig:tracker_halves}
\end{center}
%\vspace*{-5mm}
\end{figure}
%=======================

\subsection{Components}
The sensors for the SVT are $p+$-on-$n$, single sided, AC coupled, polysilicon-biased microstrip 
sensors fabricated on $<100>$ silicon and have 30 (60)~$\mu$m sense (readout) pitch over their 
$4\times10$~cm$^2$ surface. This sensor technology was selected to match the requirement of 
a $<1$\% $X_0$ per layer, single-hit resolution better than 50~$\mu$m and tolerant of a radiation 
dose of approximately $1.5\times10^{14}$~\fluenceunit{} for a six month run. The sensors 
were purchased from the Hamamatsu Photonics Corporation for the cancelled 
Run 2b upgrade of the D\O~experiment~\cite{Denisov:2001aa} which satisfied the requirement that 
the technology must be mature and available within the time and budget constraints.

@@

During the HPS Test Run, both halves were held in place by four vertical bars attached to an above rail, 
placing the front face of the crystals 147~cm from the upstream edge of the magnet and with a 
8.7~cm gap between the innermost edge of the crystals in the two halves.
{\small
\begin{figure}[]
\begin{center}
\includegraphics[width=0.45\textwidth]{figures/ECal}
%\includegraphics[width=0.45\textwidth]{figures/ecal_picture_beamline}
\caption{Rendered layout view of the ECal looking downstream.}
\label{fig:ecal}
\end{center}
\end{figure}
}
